---
title: "Cuisine and Yelp Ratings"
author: 
  - Moohaeng Sohn
thanks: "Code and data are available at: [https://github.com/alexsohn1126/yelp-analysis](https://github.com/alexsohn1126/yelp-analysis)"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(modelsummary)
library(arrow)
library(kableExtra)
library(here)
restaurants <- read_parquet(here("./data/analysis_data/restaurant_data.parquet"))
model <- readRDS(here("./models/restaurant_rating_model_logit.rds"))
```


# Introduction

Thanks to internet technology, we can get hundreds, if not thousands of reviews of restaurants around us. This means we can choose which restaurants we will go to based on those reviews. Therefore, keeping a high review rating is very important for a restaurant's long-term success.

Yelp is a website where people can post reviews about local businesses, containing 287 million reviews [@yelpfacts]. Yelp's reviews are based on the 5-star system. Users can choose between one to five stars to put on their review. These reviews are collected and averaged, which becomes the rating for the establishment. We will use a dataset from Yelp which we will dive into in @sec-data.

TODO: ADD THINGS ABOUT WHAT WE FOUND OUT AND WHY IS IT IMPORTANT

In this paper, our estimand of interest is how different factors such as the cuisine of the restaurant, or the price of the menu items in the restaurants affect the rating of a restaurant. We will first explore the dataset in @sec-data, and discuss the model the relationship between aforementioned variables using logistic regression in #sec-model. Then we will look at the results in @sec-res, finally discussing about these results in @sec-discussion.

# Data {#sec-data}

The restaurant review dataset we will use is from Yelp. Yelp offers an academic dataset for the public to use, although it is a small subset of their massive database [@yelpdataset-faq]. It only contains 

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

modelsummary(list("my model" = model))

modelplot(model, conf_level = 0.9) +
  labs(x = "90 per cent credibility interval")
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it. 



# Model {#sec-model}

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results {#sec-res}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```




# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


